---
title: "gtsummary in quarto"
format:
  html:
    theme: quartz
execute:
  echo: false
---

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

**Data Information**

The NLSY97 consists of a nationally representative sample of 8,984 men and women born during the years 1980 through 1984 and living in the United States at the time of the initial survey in 1997.

```{r}

tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```


```{r}
tableone<-tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```
```{r}
#| label: tbl-one
#| tbl-cap: "Descriptive satistics stratified by sex"
tableone

```

@tbl-one shows the descriptive statistics.

```{r}
#| echo: true
mean_sleep<-mean(nlsy$sleep_wknd, na.rm=TRUE)
mean_sleep



```

Participants received an average `r mean_sleep` hours of sleep on the weekends.

Participants received an average `r mean(nlsy$sleep_wknd, na.rm=TRUE)` hours of sleep on the weekends.

Rounded value: Participants received an average `r round(mean_sleep, 2)` hours of sleep on the weekends.

The median hours of sleep participants received on the weekend was `r median(nlsy$sleep_wknd, na.rm=TRUE)` hours.

```{r}
library(gtsummary)
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
#| output-location: slide
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, sleep_wknd, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
	missing_text = "Missing"

 ,
)|> 
  add_overall(last = TRUE)
table1

median_overall<-inline_text(table1, variable = "sleep_wknd", column = "stat_0")
```

The median hours of sleep on the weekend for the overall sample was `r median_overall` hours. 


